---
title: "Arkas: Calculate TPMs in Much Less Time"
author: "Timothy Triche Jr., Anthony Colombo, Harold Pimmentel"
output:
  html_vignette:
  toc: true
  number_sections: true
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteDepends{arkasData}
  %\VignetteIndexEntry{Arkas: Repetitive Elements Quantification In Much Less Time}
  \usepackage[utf8]{inputenc}
date: "`r format(Sys.time(), '%d %B, %Y')`"
---






```{r load_NS}
suppressPackageStartupMessages(library(arkas))
library(arkasData)
suppressPackageStartupMessages(library(TxDbLite))
samples<-c("n1","n2","n4","s1","s2","s4")
pathBase<-system.file("extdata",package="arkasData")
merged <- mergeKallisto(samples, outputPath=pathBase)
assays(merged)
kallistoVersion(merged)
transcriptomes(merged)
tail(tpm(merged))


```

#Calculating TPM
Arkas can calculate TPMs using the effective length, or an accessor tpm().
Here we only show a TPM calculation for one sample named 'n1', but it is a uniform process and can easily be applied to many samples.
For this demo we do see two outliers which an absolute difference of greater than 0.3 which is by defintion a true calculation error.  By looking at the raw values derived from calcTpm(), tpm() and the raw kallisto derived .tsv number, each of the values are very close and it appears to be a rounding issue.  The calculation was numerically precise for 213780/213782  which has high accuracy overall.

```{r error_check,fig.height=9,fig.width=7}
 n1Tsv<-read.csv(paste0(pathBase,"/n1/abundance.tsv"),header=TRUE,sep="\t")
 head(n1Tsv)
 xx2<-n1Tsv$est_counts
 leng<-n1Tsv$eff_length
 xx2<-as.matrix(xx2)
 leng<-as.matrix(leng)
 colnames(xx2)<-"n1"
 colnames(leng)<-"n1"
 test1<-calcTpm(xx2,leng)
 err<-abs(test1-n1Tsv$tpm)
 err2<-abs(tpm(merged)[,1]-n1Tsv$tpm)

 plot(err)
 title("calcTpm N1 absolute Errors")
 plot(err2)
 title("tpm() N1 absolute Errors")
```

##Comparing Outliers 
Here we look at the calculated TPM errors greater than 0.3

```{r comparison}
 rownames(err)<-n1Tsv$target_id
 highErrs<-as.matrix(err[which(err>0.3),])
 rownames(test1)<-n1Tsv$target_id
 n1Tsv[n1Tsv$target_id%in%rownames(highErrs),] #kallisto dervied

 test1[rownames(test1)%in%rownames(highErrs),] #arkas::calcTpm()
 tpp<-tpm(merged)
 tpp[rownames(tpp)%in%rownames(highErrs),1] #arkas::tpm()

```

##Calculating N2 


```{r n2_error_check,fig.height=9,fig.width=7}
 n2Tsv<-read.csv(paste0(pathBase,"/n2/abundance.tsv"),header=TRUE,sep="\t")
 head(n2Tsv)
 xx2<-n2Tsv$est_counts
 leng<-n2Tsv$eff_length
 xx2<-as.matrix(xx2)
 leng<-as.matrix(leng)
 colnames(xx2)<-"n2"
 colnames(leng)<-"n2"
 test1<-calcTpm(xx2,leng)
 err<-abs(test1-n2Tsv$tpm)
 err2<-abs(tpm(merged)[,2]-n2Tsv$tpm)

 plot(err)
 title("calcTpm N2 absolute Errors")
 plot(err2)
 title("tpm() N2 absolute Errors")
```

##Comparing Outliers For N2
Here we look at the calculated TPM errors greater than 0.3

```{r n2_comparison}
 rownames(err)<-n2Tsv$target_id
 highErrs<-as.matrix(err[which(err>0.3),])
 rownames(test1)<-n2Tsv$target_id
 n2Tsv[n2Tsv$target_id%in%rownames(highErrs),] #kallisto dervied

 test1[rownames(test1)%in%rownames(highErrs),] #arkas::calcTpm()
 tpp<-tpm(merged)
 tpp[rownames(tpp)%in%rownames(highErrs),2] #arkas::tpm()

```

##For N4

```{r n4_error_check,fig.height=9,fig.width=7}
 n4Tsv<-read.csv(paste0(pathBase,"/n4/abundance.tsv"),header=TRUE,sep="\t")
 head(n4Tsv)
 xx2<-n4Tsv$est_counts
 leng<-n4Tsv$eff_length
 xx2<-as.matrix(xx2)
 leng<-as.matrix(leng)
 colnames(xx2)<-"n4"
 colnames(leng)<-"n4"
 test1<-calcTpm(xx2,leng)
 err<-abs(test1-n4Tsv$tpm)
 err2<-abs(tpm(merged)[,3]-n4Tsv$tpm)

 plot(err)
 title("calcTpm N4 absolute Errors")
 plot(err2)
 title("tpm() N4 absolute Errors")
```

##Comparing Outliers For N4
Here we look at the calculated TPM errors greater than 0.3

```{r n4_comparison}
 rownames(err)<-n4Tsv$target_id
 highErrs<-as.matrix(err[which(err>0.3),])
 rownames(test1)<-n4Tsv$target_id
 n4Tsv[n4Tsv$target_id%in%rownames(highErrs),] #kallisto dervied

 test1[rownames(test1)%in%rownames(highErrs),] #arkas::calcTpm()
 tpp<-tpm(merged)
 tpp[rownames(tpp)%in%rownames(highErrs),3] #arkas::tpm()

```

##S1 Error Checking


```{r s1_error_check,fig.height=9,fig.width=7}
 s1Tsv<-read.csv(paste0(pathBase,"/s1/abundance.tsv"),header=TRUE,sep="\t")
 head(s1Tsv)
 xx2<-s1Tsv$est_counts
 leng<-s1Tsv$eff_length
 xx2<-as.matrix(xx2)
 leng<-as.matrix(leng)
 colnames(xx2)<-"s1"
 colnames(leng)<-"s1"
 test1<-calcTpm(xx2,leng)
 err<-abs(test1-s1Tsv$tpm)
 err2<-abs(tpm(merged)[,4]-s1Tsv$tpm)

 plot(err)
 title("calcTpm S1 absolute Errors")
 plot(err2)
 title("tpm() S1 absolute Errors")
```
 
##Comparing Outliers 
Here we look at the calculated TPM errors greater than 0.3

```{r s1_comparison}
 rownames(err)<-s1Tsv$target_id
 highErrs<-as.matrix(err[which(err>0.3),])
 rownames(test1)<-s1Tsv$target_id
 s1Tsv[s1Tsv$target_id%in%rownames(highErrs),] #kallisto dervied

 test1[rownames(test1)%in%rownames(highErrs),] #arkas::calcTpm()
 tpp<-tpm(merged)
 tpp[rownames(tpp)%in%rownames(highErrs),4] #arkas::tpm()
 
```

##S2 Error Checking

```{r s2_error_check,fig.height=9,fig.width=7}
 s2Tsv<-read.csv(paste0(pathBase,"/s2/abundance.tsv"),header=TRUE,sep="\t")
 head(s2Tsv)
 xx2<-s2Tsv$est_counts
 leng<-s2Tsv$eff_length
 xx2<-as.matrix(xx2)
 leng<-as.matrix(leng)
 colnames(xx2)<-"s2"
 colnames(leng)<-"s2"
 test1<-calcTpm(xx2,leng)
 err<-abs(test1-s2Tsv$tpm)
 err2<-abs(tpm(merged)[,5]-n2Tsv$tpm)
 
 plot(err)
 title("calcTpm S2 absolute Errors")
 plot(err2)
 title("tpm() S2 absolute Errors")
```

##Comparing Outliers For S2
Here we look at the calculated TPM errors greater than 0.3

```{r s2_comparison}
 rownames(err)<-s2Tsv$target_id
 highErrs<-as.matrix(err[which(err>0.3),])
 rownames(test1)<-s2Tsv$target_id
 s2Tsv[s2Tsv$target_id%in%rownames(highErrs),] #kallisto dervied
 
 test1[rownames(test1)%in%rownames(highErrs),] #arkas::calcTpm()
 tpp<-tpm(merged)
 tpp[rownames(tpp)%in%rownames(highErrs),5] #arkas::tpm()

```


##S4 Error Checking

```{r s4_error_check,fig.height=9,fig.width=7}
 s4Tsv<-read.csv(paste0(pathBase,"/s4/abundance.tsv"),header=TRUE,sep="\t")
 head(s4Tsv)
 xx2<-s4Tsv$est_counts
 leng<-s4Tsv$eff_length
 xx2<-as.matrix(xx2)
 leng<-as.matrix(leng)
 colnames(xx2)<-"s4"
 colnames(leng)<-"s4"
 test1<-calcTpm(xx2,leng)
 err<-abs(test1-s4Tsv$tpm)
 err2<-abs(tpm(merged)[,6]-s4Tsv$tpm)
 
 plot(err)
 title("calcTpm S4 absolute Errors")
 plot(err2)
 title("tpm() S4 absolute Errors")
```

##Comparing Outliers For S4
Here we look at the calculated TPM errors greater than 0.3

```{r s4_comparison}
 rownames(err)<-s4Tsv$target_id
 highErrs<-as.matrix(err[which(err>0.3),])
 rownames(test1)<-s4Tsv$target_id
 s4Tsv[s4Tsv$target_id%in%rownames(highErrs),] #kallisto dervied
 
 test1[rownames(test1)%in%rownames(highErrs),] #arkas::calcTpm()
 tpp<-tpm(merged)
 tpp[rownames(tpp)%in%rownames(highErrs),6] #arkas::tpm()

```

